EldersAIR: data merging for analysis dataset

### 2021-03-31 manuscript_tables_figures.Rmd
tables and initial model for primary analysis

elders_pm <- read_rds(paste0(file_path, "Output/elders_pm_clean.rds")) 
```

@@

>>>>

treatments_blind_nn <- read_xlsx(paste0(file_path, "Input/NN/treatments_blind.xlsx")) %>% 
  mutate(area = "NN")

treatments_blind_npt <- read_xlsx(paste0(file_path, "Input/NP/treatments_blind.xlsx")) %>% 
  mutate(area = "NPT")

treatments_blind <- rbind(treatments_blind_nn, treatments_blind_npt) %>% 
  mutate(home_id_num = as.factor(HomeID),
         treatment_blind = as.factor(CodedCondition),
         area = as.factor(area)) %>% 
  select(area, home_id_num, treatment_blind)

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds")) %>% 
  left_join(treatments_blind, by = c("area", "home_id_num")) 
```

@@

```{r}
# summary data for manuscript tables
char_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>%
  arrange(home, winter_id, sampling_visit) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(treatment_blind) %>% 
  mutate(treatment_n = n()) %>% 
  group_by(treatment_blind, new_var) %>% 
  mutate(n = n(),
         percent = n()/treatment_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(treatment_blind, new_var)
demographics_summary
}

# distinct by participant
char_funct_total("gender")
char_funct_treatment("gender")
char_funct_total("hispanic")
char_funct_treatment("hispanic")
char_funct_total("race")
char_funct_treatment("race")
char_funct_total("education")
char_funct_treatment("education")

#distinct by home
char_funct_total("residents_over55")
char_funct_treatment("residents_over55")
char_funct_total("income_3level")
char_funct_treatment("income_3level")
char_funct_total("home_floors_2level")
char_funct_treatment("home_floors_2level")
char_funct_total("home_sqm_2level")
char_funct_treatment("home_sqm_2level")
char_funct_total("home_year_built_2level")
char_funct_treatment("home_year_built_2level")
char_funct_total("home_pets_3level")
char_funct_treatment("home_pets_3level")
char_funct_total("residents_smoke")
char_funct_treatment("residents_smoke")
char_funct_total("residents_smoke_inside")
char_funct_treatment("residents_smoke_inside")

char_funct_total("stove_age")
char_funct_treatment("stove_age")
char_funct_total("chimney_clean")
char_funct_treatment("chimney_clean")
char_funct_total("wood_collect_method")
char_funct_treatment("wood_collect_method")
char_funct_total("wood_collect")
char_funct_treatment("wood_collect")
char_funct_total("stove_grade")
char_funct_treatment("stove_grade")
char_funct_total("burn_level")
char_funct_treatment("burn_level")
```

```{r}
# summary data for manuscript tables - 1 response from Winter 1
num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  filter(winter_id == 1) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  distinct(adult_id_char, sampling_visit, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(new_var)) %>% 
  dplyr::select(new_var) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>%
  filter(winter_id == 1) %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  distinct(adult_id_char, sampling_visit, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(new_var)) %>% 
  dplyr::select(treatment_blind, new_var) %>% 
  group_by(treatment_blind) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
demographics_summary
}

# distinct by participant
num_funct_total("age")
num_funct_treatment("age")

num_funct_total("sum_act_smoking")
num_funct_treatment("sum_act_smoking")
num_funct_total("sum_act_windows")
num_funct_treatment("sum_act_windows")
num_funct_total("sum_act_sweep")
num_funct_treatment("sum_act_sweep")
num_funct_total("sum_act_door")
num_funct_treatment("sum_act_door")

# distinct by home
num_funct_total("home_bedrooms")
num_funct_treatment("home_bedrooms")

# summary data for manuscript tables - averaged over multiple Winter 1 visits
num_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(home, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(mean_var) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

num_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>%
  mutate(new_var = as.numeric(new_var)) %>% 
  group_by(home, winter_id) %>% 
  mutate(mean_var = mean(new_var, na.rm = T)) %>% 
  ungroup() %>% 
  filter(winter_id == 1) %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  filter(!is.na(mean_var)) %>% 
  dplyr::select(treatment_blind, mean_var) %>% 
  group_by(treatment_blind) %>% 
  summarize(mean_new_var = mean(mean_var), sd_var = sd(mean_var), n_var = n(), 
            min_var = min(mean_var), med_var = median(mean_var), max_var = max(mean_var))
demographics_summary
}

# distinct by participant
num_funct_total("bp_sys")
num_funct_treatment("bp_sys")
num_funct_total("bp_dia")
num_funct_treatment("bp_dia")
num_funct_total("bmi")
num_funct_treatment("bmi")
num_funct_total("fvc")
num_funct_treatment("fvc")
num_funct_total("fev1")
num_funct_treatment("fev1")
num_funct_total("fev1_fvc_ratio")
num_funct_treatment("fev1_fvc_ratio")
num_funct_total("fvc_pp")
num_funct_treatment("fvc_pp")
num_funct_total("fev1_pp")
num_funct_treatment("fev1_pp")

num_funct_total("home_act_smoking")
num_funct_treatment("home_act_smoking")
num_funct_total("home_act_windows")
num_funct_treatment("home_act_windows")
num_funct_total("home_act_sweep")
num_funct_treatment("home_act_sweep")
num_funct_total("home_act_door")
num_funct_treatment("home_act_door")

# distinct by home
num_funct_total("pm_mean_winter")
num_funct_treatment("pm_mean_winter")
```

### 2021-03-31 merge_data_files.Rmd
tables and initial model for primary analysis

# summary(health_pm_visit)
```

Save final dataset
```{r}
write_rds(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))

write_csv(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")

# save variable names to use in data dictionary
save_var_names <- data.frame(t(health_pm_visit)) %>% 
  rownames_to_column() %>% 
  select(rowname)

>>>>

elders_cleaned_dataset_by_visit <- health_pm_visit %>% 
  mutate(home_floors = as.factor(home_floors),
         home_floors_2level = fct_collapse(home_floors,
                                           "1" = "1",
                                           "2+" = c("2", "3")),
         home_sqm = home_sqft*0.0929,
         home_sqm_2level = cut(home_sqm, breaks = c(0, 112, 10000),
                               labels = c("<111", "111+")),
         home_bedrooms = as.numeric(home_bedrooms),
         home_year_built_2level = cut(home_year_built, 
                                      breaks = c(0, 1981, 10000),
                                      labels = c("<1981", "1981+")),
         home_pets = as.factor(home_pets),
         home_pets_3level = fct_collapse(home_pets,
                                         "0" = "0",
                                         "1" = "1",
                                         "2+" = c("2", "3", "4", "5", "6", "7", "10")),
         income = as.factor(income),
         income_3level = fct_collapse(income,
                                      "< $20,000" = "< $20,000",
                                      "$20,000 to $39,999" = c("$20,000 to $29,999", "$30,000 to $39,999"),
                                      "$40,000+" = c("$40,000 to $49,999", "$50,000 to $74,999", 
                                                     "$75,000 to $99,999", ">= $100,000")),
         stove_age = factor(stove_age,
                            levels = c("0-5 years old", "6-10 years old", "11-15 years old",
                                       "16 + years old", "Do not know")),
         chimney_clean = factor(chimney_clean,
                            levels = c("Less than 6 months ago", "6-12 months ago", 
                                       "12-18 months ago", "More than 18 months ago")),
         burn_level = factor(burn_level,
                            levels = c("No burning", "Light burning", 
                                       "Average burning", "Heavy burning")),
         wood_collect = factor(wood_collect,
                            levels = c("Less than 1 week", "Between 1 week and 1 month", 
                                       "Between 1 month and 3 months", "Between 3 months and 6 months",
                                       "Between 6 months and 1 year", "More than 1 year"))) %>% 
  group_by(adult_id_char, winter_id) %>% 
  mutate(sum_act_smoking = sum(home_act_smoking, na.rm = T),
         sum_act_candle = sum(home_act_candle, na.rm = T),
         sum_act_door = sum(home_act_door, na.rm = T),
         sum_act_gas = sum(home_act_gas, na.rm = T),
         sum_act_elect = sum(home_act_elect, na.rm = T),
         sum_act_propane = sum(home_act_propane, na.rm = T),
         sum_act_oil = sum(home_act_oil, na.rm = T),
         sum_act_incense = sum(home_act_incense, na.rm = T),
         sum_act_kerosene = sum(home_act_kerosene, na.rm = T),
         sum_act_windows = sum(home_act_windows, na.rm = T),
         sum_act_sweep = sum(home_act_sweep, na.rm = T)) %>% 
  group_by(adult_id_char) %>% 
  arrange(adult_id_char, sampling_date) %>% 
  mutate(first_date = first(sampling_date),
         cohort = round_date(first_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  mutate(cohort = as.factor(cohort)) %>% 
  select(-trash)

elders_distinct <- elders_cleaned_dataset_by_visit %>% 
  arrange(area, home, winter_id, sampling_visit) %>% 
  distinct(home, .keep_all = T)

median(elders_distinct$home_year_built, na.rm = T)
table(elders_distinct$sum_act_candle)
```

Save final dataset
```{r}
write_rds(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))

write_csv(elders_cleaned_dataset_by_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")

# save variable names to use in data dictionary
save_var_names <- data.frame(t(elders_cleaned_dataset_by_visit)) %>% 
  rownames_to_column() %>% 
  select(rowname)
